### 2022-05-17 main/chapter_2ljets.tex
finish variables cheatsheet

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
for which the lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\etmiss$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $t\bar t$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

\paragraph{My contributions}
\begin{itemize}
\item Design, implementation and execution of the electroweak part of the analysis.
\item Analysis Contact from January to October 2020.
\item Production of main data inputs for all three analyses
and the RJR $3\ell$ search \TODO{cite}.
I produced all systematic variations for the background samples and all
electroweak signal samples.
Other group members assisted with the central background sample.
\item Integration with ATLAS combinations and pMSSM scan efforts by
performing their validation studies and serializing the analysis results.
\item Preparation of electroweak results for publication in the paper, and the
public HEPData record~\cite{maguire2017hepdata}.
\end{itemize}

@@

Uncomfortably many event variables are used to define the regions of this
analysis.
Definitions of these variables are collected here.
All variables with a double object subscript (such as $\mll$, $\rjj$) refer to
the first two of those particles when ordered in decreasing $\pt$.
\begin{itemize}
\item $\etmiss$: missing transverse momentum.
\item $\etmisssig$: object-based missing transverse momentum significance.
It scales $\etmiss$ by an uncertainty derived from the resolutions of
reconstructed objects~\cite{atlas_met_significance}.
\item $\mll$: invariant mass of the hardest two leptons.
\item $\mjj$: invariant mass of the two hardest jets.
\item $\mbb$: invariant mass of the two hardest b-tagged jets.
\item $\mjetone$: mass of the hardest jet.
% TODO
% \item $\Rll$: estimated angular distance in $\eta,\phi$ between the lepton pair.
% \item $\Rjj$: estimated angular distance in $\eta,\phi$ between the two hardest jets.
% \item $\Delta\phi(p_{\ell\ell}, \Etmiss)$: estimated distance in azimuthal
% angle between the vector sum of the lepton pair and the missing transverse momentum.
% \item $\mttwo(\ell, \ell; 0)$: the `stransverse mass' between the two
% leptons and the missing transverse momentum.
% If two massive particles are pair produced and each decays to a
% lepton and something massless and invisible,
% then this estimates a greatest lower bound on the invariant
% mass of the pair produced particles.
% $\mttwo(\ell, \ell; 0)$ is useful for excluding standard model backgrounds;
% di-leptonic ttbar ($tt\rightarrow WW\rightarrow\ell\nu\ell\nu$)
% and $WW\rightarrow\ell\nu\ell\nu$ have a kinematic endpoint
% at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.
% \item $\nbtag$: number of b-tagged jets with $\pt > 20$ GeV,
% where the tag is taken on the mv2c10 classifier at the $77\%$ working point.
% \item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
% \item $\pt(j_1)$: transverse momentum of the hardest jet.
% \item $\Delta\phi(\mathrm{jet}_1,\etmiss)$: azimuthal angle between
% the hardest jet and missing transverse momentum vector estimates.
\end{itemize}
Most of these are physical properties of particles, or collection of particles.
Clearly, the stated values are not exact, but estimated values based on
approximate reconstructions.
While I think this distinction is important to remember, I have not spelled it
out on every line.
Brevity also has value.

@@

Splitting regions on $\etmisssig$ is the core idea behind the design.
The basic motivation for this design are displayed in
Figure~\ref{fig:2ljets_presel_met};
$\etmisssig$ does better than $\etmiss$ in separating background and signal
contributions.
But there is more --- $\etmisssig$ also does a decent job of separating signal
models from each other;
signals with larger mass differences in their decays to invisibles tend to have
larger $\etmisssig$.
Binning regions in $\etmisssig$ therefore also improves sensitivity to
individual parameter points and allows targeted selections on other, locally-relevant,
event variables.

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\etmisssig$ beats $\etmiss$ alone in sensitivity to
example signal models.
At large $\etmisssig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\etmiss$.
\\[0.5em]
This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
signal leptons with
$\mll \in (71, 111)~\eV[G]$,
$\mjj \in (60, 110)~\eV[G]$,
$\nbtag \leq 1$,
$\etmiss > 100~\eV[G]$, and
$\etmisssig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

>>>>

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
for which the lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\met$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $t\bar t$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

\paragraph{My contributions}
\begin{itemize}
\item Design, implementation and execution of the electroweak part of the analysis.
\item Service as `analysis contact' from January to October 2020.
\item Production of main data inputs for all three analyses
and the RJR $3\ell$ search \TODO{cite}.
Production of all systematic variations for the background samples;
other group members assisted with the central background sample.
Production of all electroweak signal samples and their systematic variations.
\item Integration with ATLAS combinations and pMSSM scan efforts by
performing their validation studies and serializing the analysis results.
\item Preparation of electroweak results for publication in the paper \TODO{cite} and the
public HEPData record~\cite{maguire2017hepdata}.
\end{itemize}

@@

\subsection{$\met$ significance}
\label{sec:metsig}
It scales $\met$ by an uncertainty derived from the resolutions of
reconstructed objects~\cite{atlas_met_significance}.

\subsection{Stranverse mass $\mttwo$}
\label{sec:mt2}
% angle between the vector sum of the lepton pair and the missing transverse momentum.
% leptons and the missing transverse momentum.
% If two massive particles are pair produced and each decays to a
% lepton and something massless and invisible,
% then this estimates a greatest lower bound on the invariant
% mass of the pair produced particles.
% $\mttwo(\ell, \ell; 0)$ is useful for excluding standard model backgrounds;
% di-leptonic ttbar ($tt\rightarrow WW\rightarrow\ell\nu\ell\nu$)
% and $WW\rightarrow\ell\nu\ell\nu$ have a kinematic endpoint
% at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.

\subsection{b-tagging}
\label{sec:btagging}
% with $\pt > 20$ GeV,
% where the tag is taken on the mv2c10 classifier at the $77\%$ working point.

\subsection{Cheat-sheet}
Uncomfortably many event variables are used to define the regions of this
analysis.
To aid future revision of their meanings, definitions and references for these
variables are collected here.
\begin{itemize}
\item $\met$: missing transverse momentum
\item $\metsig$: object-based missing transverse momentum significance;\\
described in Section~\ref{sec:metsig}
\item $\ptjone$: transverse momentum of the hardest jet%
\vspace{0.5em}
\item $\mll$: invariant mass of the hardest two leptons
\item $\mjj$: invariant mass of the two hardest jets
\item $\mbb$: invariant mass of the two hardest b-tagged jets
\item $\mjetone$: mass of the hardest jet
\item $\mttwoll$: `stransverse` mass of the lepton pair with massless invisibles;\\
described in Section~\ref{sec:mt2}
\vspace{0.5em}
\item $\rll$: distance in $\eta\textrm{--}\phi$ between the lepton pair
\item $\rjj$: distance in $\eta\textrm{--}\phi$ between the two hardest jets%
\vspace{0.5em}
\item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
\item $\nbtag$: number of b-tagged jets;
described in Section~\ref{sec:btagging}
\vspace{0.5em}
\item $\dphillmet$: azimuthal angle between the hardest jet and $\ptmiss$
\item $\Delta\phi(\mathrm{jet}_1,\met)$: azimuthal angle between
the hardest jet and $\ptmiss$
\end{itemize}
All of these are physical properties of particles or collections of particles.
Please remember that none is an exact or true value;
all are are estimates based on approximate reconstructions.
While this distinction is easy to drop in lax language, is necessary to
understand the shapes of noisy variables such as $\mjj$ and indeed the
existence of $\metsig$.

@@

Splitting regions on $\metsig$ is the core idea behind the design.
The basic motivation for this design are displayed in
Figure~\ref{fig:2ljets_presel_met};
$\metsig$ does better than $\met$ in separating background and signal
contributions.
But there is more --- $\metsig$ also does a decent job of separating signal
models from each other;
signals with larger mass differences in their decays to invisibles tend to have
larger $\metsig$.
Binning regions in $\metsig$ therefore also improves sensitivity to
individual parameter points and allows targeted selections on other, locally-relevant,
event variables.

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\metsig$ beats $\met$ alone in sensitivity to
example signal models.
At large $\metsig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\met$.
\\[0.5em]
This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
signal leptons with
$\mll \in (71, 111)~\eV[G]$,
$\mjj \in (60, 110)~\eV[G]$,
$\nbtag \leq 1$,
$\met > 100~\eV[G]$, and
$\metsig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}
